Fast Automatic Bayesian Cubature Using Lattice Sampling

### 2019-07-22 BayesianCubaturePaper2018 - Rev1b.tex
polished

The process outlined above typically requires a computational cost of $O(N_{\text{opt}}n^3)$, where $N_{\text{opt}}$ is the number of optimization steps required to identify the hyperparameters. Our innovation is to pair low discrepancy nodes with matching kernels that lower the  computational cost to $O(N_{\text{opt}} n \log n)$.   This approach is demonstrated using rank-1 lattice sequences and shift-invariant kernels.  Our algorithm is  implemented in the Guaranteed Automatic Integration Library (GAIL).

@@

Section \ref{sec:BC} explains the Bayesian approach to estimating the posterior cubature error and defines our automatic Bayesian cubature. Although much of this material is known, it is included for completeness.  We end Section \ref{sec:BC}  by demonstrating why Bayesian cubature is typically computationally expensive.
Section \ref{sec:fast_BC}  introduces the concept of covariance kernels that match the nodes, which expedites the computations required by our automatic Bayesian cubature. 
Section \ref{sec:shift_invariant_kernel} implements this concept for shift-invariant kernels and rank-1 lattice nodes.  This section also describes how to avoid cancellation error for kernels of product form.  Numerical examples are provided in Section \ref{sec:NumExp} to demonstrate our new algorithm.  We conclude with a brief discussion.

@@

We use the following parameter values in the simulation: 
\begin{align*}
d' = 3, \quad \va = \begin{pmatrix}
-6 \\ -2 \\ -2
\end{pmatrix}, \quad 
\vb = \begin{pmatrix}
5 \\ 2 \\ 1
\end{pmatrix} , \quad 
\mL = \begin{pmatrix}
4 & 1 & 1 \\ 0 & 1 & 0.5 \\ 0 & 0 & 0.25
\end{pmatrix}.
\end{align*}
The node sets are randomly scrambled Sobol' points \cite{DicEtal14a,DicPil10a}.  The results are for 400 randomly chosen $\varepsilon$ in the interval $[10^{-5}, 10^{-2}]$ as shown in \figref{fig:MVN_Metern_d2b2}. In each run, the nodes are randomly scrambled.  We  observe the algorithm meets the error criterion 95\% of the time even though we used 99\% credible interval. One possible explanation is, the matrix inversions in the algorithm are ill-conditioned leading to numerical inaccuracies.
%with Intel i7 3630QM and 16GB RAM memory
%On our test computer, 
As shown in \figref{fig:MVN_Metern_d2b2}, computation time increases rapidly with $n$. 
The  maximum likelihood estimation of $\vtheta$, which requires repeated evaluation of the objective function, is the most time consuming of all. It takes tens of seconds to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$.   In contrast, this example in Section \ref{sec:NumExp} take less than a hundredth of a second to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$ using our new algorithm. Not only is the Bayesian cubature with the Mat\'ern kernel slow, but also $\mC_\vtheta$ becomes highly ill-conditioned as $n$ increases.
So, Algorithm \ref{algorithm1}  in its current form is impractical when $n$ must be large. 
%\FJHNote{Jags, please check this re-write} %\JRNote{looks good}

\begin{figure}
	\centering
		\includegraphics[width=0.95\linewidth]{MVN_guaranteed_time_Matern_d2_2019-Jun-29}
	\centering
		\includegraphics[width=0.95\linewidth]{MVN_rapid_n_vs_time_Matern_d2_2019-Jun-29}
	\caption{Multivariate Gaussian probability estimated using Mat\'ern kernel with $d=2$ using empirical stopping criterion. Top: Guaranteed integration within error tolerance $\varepsilon$. Bottom: Computation time rapidly increases with increase of $n$. These figures can be reproduced using \code{matern\_guaranteed\_plots.m} in GAIL.}
	\label{fig:MVN_Metern_d2b2}
\end{figure}

@@

The generic automatic Bayesian cubature algorithm described in the previous section requires $\Order\bigl(n \$(f) + \break N_{\opt} [n^2 \$(C_\vtheta)  + n^3]\bigr)$ operations to estimate $\vtheta$, compute the credible interval width, and compute the cubature. Now we explain how to speed up the calculations. A key is to choose covariance kernels that match the nodes, $\{\vx_i\}_{i=1}^n$, so that the vector-matrix operations required by Bayesian cubature can be accomplished using fast transforms at a cost of $\Order\bigl(n \$(f) + N_{\opt} [n \$(C_\vtheta)  + n \log(n)] \bigr)$.

@@

We make three assumptions that allow the fast computation:
\begin{subequations} \label{fastcompAssump}
	\begin{gather}
	\label{fastcompAssumpA}
	\mV \text{ may be identified analytically}, \\
	\label{fastcompAssumpB}
	\vv_1 = \vV_1 = \vone, \\
	\label{fastcompAssumpC}
	\mV^H \vb  \text{ requires only $\Order(n \log(n))$ operations } \forall \vb.
	\end{gather}
\end{subequations}
We call the transformation $\vb \mapsto \mV^H \vb$ a \emph{fast transform} and $C$ a \emph{fast transform kernel}.

@@

\begin{theorem} \label{thm:fastparam}
Under assumptions \eqref{fastcompAssump}, the parameters and credible interval widths in Theorem \ref{thm:param} may be expressed in terms of the fast transforms of the function data, the first column of the Gram matrix, and $\vc_\vtheta$ as follows:
\begin{align}
\nonumber
m_\MLE &=  m_{\full} = m_{\GCV} =  \frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\\
\nonumber
s^2_\MLE 
& =
\frac{1}{n^2} 
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}, \\
\nonumber
\widehat{\sigma}^2_{\textup{full}}
& =
\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}
\\
\nonumber
& \qquad \times
\left[\frac{\lambda_1}{n}{\left(1 - \frac{\widetilde{c}_1}{\lambda_1}\right)^2} + \left(c_0  - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i}\right) \right], \\
\nonumber 
s^2_{\textup{GCV}} & =  \frac 1{n} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1},
\end{align}
\begin{subequations}
	\label{fastTheta}
	\begin{align}
\nonumber
\vtheta_\MLE
&= 
\argmin_{\vtheta}
\left[
\log\left(
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta i}}
\right) \right .\\ 
\label{eqn_MLE_loss_func_optimized_2} %
& \hspace{3cm} \left .
 + \frac{1}{n}\sum_{i=1}^n \log(\lambda_{\vtheta i})
\right],\\
\nonumber 
\vtheta_{\GCV} 
&= \argmin_\vtheta \left[ \log \left ( \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta i}^2} 
\right) \right .\\
\label{thetaGCV}
& \qquad \qquad \qquad \qquad  \left . -2\log\left( \sum_{i=1}^n \frac{1}{\lambda_{\vtheta i}} \right)
\right], 
\end{align}
\end{subequations}
\begin{align}
\nonumber
\hmu_\MLE  &= \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} +
\frac 1n \sum_{i=2}^n \frac{ \widetilde{c}_i^* \widetilde{y}_i}{\lambda_i}, \\
\nonumber
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
}, \\
\nonumber
\err_{\full} & = t_{n-1,0.995} \hsigma_{\textup{full}}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} \right.
\\ 
\nonumber
&\qquad \qquad \left . \times
\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
\right\}^{1/2}.
\end{align}
Under the further assumption \eqref{addAssump}, it follows that 
\begin{equation}
\label{muhatGCV-FB-MLE-Simple}
\hmu_\MLE  = \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\end{equation}
and so $\hmu$ is simply the sample mean.  Also, under assumption \eqref{addAssump}, the credible interval widths simplify to
\begin{align}
\label{eq:errMLEAllAsump}
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( 1 -  \frac{n}{\lambda_1} \right) 
}, \\
\label{FJH:eq:errFullSimple}
\err_{\textup{full}}
&=
t_{n-1,0.995}
\sqrt{\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  \left(\frac{\lambda_1}{n}  - 1  \right)}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} 
\right. \\ &\qquad \qquad \left . \times
\left( 1 -  \frac{n}{\lambda_1} \right)  
\right\}^{1/2}. \label{errGCVSimple}
\end{align}
In the formulas for the credible interval widths, $\vlambda$ depends on $\vtheta$, and  $\vtheta$ is assumed to take on the values $\vtheta_{\MLE}$ or $\vtheta_{\GCV}$ as appropriate.
\end{theorem}

@@

The preceding sections lay out an automatic Bayesian cubature algorithm whose computational cost is drastically reduced.  However, this algorithm relies on covariance kernel functions, $C_\vtheta$ and node sets, $\{\vx_i\}_{i=1}^n$ that satisfy assumptions \eqref{fastcompAssump}.  We  also want to satisfy assumption \eqref{addAssump}.  
%To facilitate the fast transform, it is assumed in this section and the next that $n$ is power of $2$.  
To facilitate the fast transform, $n$ must be power of $2$.

@@

The set of nodes used is defined by a shifted extensible integration lattice node sequence, which takes the form
\begin{equation} \label{eqn:lattice_def}
\vx_{i} = \vh \phi(i-1) + \vDelta \mod \vone, \qquad i \in \naturals.
\end{equation} 
Here, $\vh$ is a $d$-dimensional generating vector of positive integers, $\vDelta$ is some point in $[0,1)^d$, often chosen at random, and $\{\phi(i)\}_{i=0}^n$ is the van der Corput sequence, defined by reflecting the binary digits of the integer about the decimal point, i.e., 
\begin{align} \label{vdCDef}
\begin{array}{r|ccccccccccccc}
i & 0 & 1 & 2 & 3 & 4 &  5 & 6 & 7 & \cdots \\
i & 0_2 & 1_2 & 10_2 & 11_2 & 100_2 & 101_2 & 110_2 & 111_2  & \cdots\\
\toprule
\phi(i) & {}_2.0 &  {}_2.1 & {}_2.01 &  {}_2.11  & {}_2.001 &  {}_2.101 & {}_2.011 &  {}_2.111 & \cdots\\
\phi(i) & 0 &  0.5 &  0.25 & 0.75 &  0.125 & 0.625  &  0.375 & 0.875 & \cdots
\end{array}
\end{align}
Note that 
\begin{multline} \label{phiprop}
\phi:\{0, \ldots, n-1 \} \to \{0, 1/n, \ldots, (n-1)/n\} \\
\text{is one-to-one},
\end{multline}
assuming $n$ is a power of $2$.

An example of $64$ nodes is given in \figref{latticefig}.  The even coverage of the unit cube is ensured by a well chosen generating vector.  The choice of generating vector is typically done offline by computer search.  See \cite{DicEtal14a,HicNie03a} for more on extensible integration lattices.
\begin{figure}[htp]
	\centering
	\includegraphics[height=5cm]{ShiftedLatticePoints}
	\caption{Example of a shifted integration lattice node set  in $d=2$. 
	This figure can be reproduced using \code{PlotPoints.m} in GAIL} \label{latticefig}
\end{figure}

\subsection{Shift Invariant Kernels}
The covariance functions $C_\vtheta$ that match integration lattice node sets have the form
\begin{align} \label{eq:shInv}
C_\vtheta(\vt,\vx) = K_\vtheta(\vt - \vx \bmod \vone).
\end{align}
This is called a \emph{shift-invariant kernel} because shifting both arguments of the covariance function by the same amount leaves the value unchanged.   Of course, $K_\vtheta$ must also be of the form that ensures that $C_\vtheta$ is symmetric and positive definite, as assumed in \eqref{FJH:eq:CondPosDef}.

A family of shift-invariant kernels is constructed via even degree Bernoulli polynomials:
\begin{multline}
\label{the_kernel_eqn_bernoulli}
C_\vtheta(\vt, \vx) =
\prod_{l=1}^d \biggl[
1 - (-1)^{r} \eta B_{2r}( |{x_l-t_l}| ) \biggr], \\  
\forall \vt,\vx \in [0,1]^d, \  \vtheta = (r,\eta), \ r \in \naturals, \ \eta > 0.
\end{multline}
Symmetric, periodic, positive definite kernels of this form appear in  \cite{DicEtal14a,Hic96a}.  Bernoulli polynomials are described in \cite[Chapter 24]{OlvEtal10a}.

Larger $r$ implies a greater degree of smoothness of the covariance kernel.  Larger $\eta$ implies greater fluctuations of the output with respect to the input.  Plots of $C(\cdot, 0.3)$ are given in \figref{fig:fourierkernel-dim1} for various $r$ and $\eta$ values.

Integration lattice cubature rules are known to have convergence rates that depend on the smoothness of the integrands \cite{DicEtal14a}.  Thus, we expect integration lattice node sets to perform well regardless of the smoothness of the covariance kernel.  The bigger concern is whether the derivatives of the integrand are as smooth as the covariance kernel implies.  This topic is touched upon again in Section  \ref{period_var_tx}.

@@

\subsection{The Gram Matrix as the Permutation of a Circulant Matrix}
For general shift-invariant covariance functions, the Gram matrix takes the form of a permutation of the rows and columns of a circulant matrix. By the properties of $\phi$ in \eqref{phiprop}, it follows that
\begin{equation} \label{PermMat}
    \mP = \bigl( \delta_{\phi(i-1), (j-1)/n}  \bigr)_{i,j=1}^n
\end{equation}
is a permutation matrix, where $\delta$ is the Kronecker delta function.  Then,
\begin{align}
\nonumber
\mC_\vtheta &= \bigl ( C_\vtheta(\vx_i, \vx_j) \bigr)_{i, j = 1}^n \\
\nonumber
& = \Bigl ( K_\vtheta \bigl(\vh(\phi(i-1) - \phi(j-1) \bigr) \bmod \vone ) \Bigr)_{i, j = 1}^n \\
\nonumber
& \qquad \qquad  \text{by \eqref{eqn:lattice_def} and \eqref{eq:shInv}}  \\
& = \mP \mK_{\vtheta}  \mP^T,  \label{shInvKernGramMatrix} \\
\nonumber 
\text{where } \mK_{\vtheta} &= \bigl ( K_\vtheta \bigl (\vh (i-j)/n \bmod \vone \bigr ) \bigr)_{i, j = 1}^n.
\end{align}
\FJHNote{Check this.}

@@

Assumptions \eqref{fastcompAssumpA} and \eqref{fastcompAssumpB} can be verified by \eqref{PermMat}, \eqref{Keig}, and \eqref{Clateig}.  Assumption \eqref{fastcompAssumpC} is satisfied because $\mV^H \vb = \mW^H \mP^T \vb$ is just the discrete Fourier transform of a vectors whose rows have been permuted.  This can be performed in $\Order(n \log(n))$ operations by the FFT algorithm. A proper scaling of the kernel $K_\vtheta$ ensures that assumption \eqref{addAssump} is satisfied.

@@

Any shift-invariant covariance kernel can be written as $C_\vtheta = 1 + \rC_\vtheta$. The associated Gram matrix for $\rC_\vtheta$ is then $\rmC_\vtheta = \mC_\vtheta - \vone \vone^T$, and the eigenvalues of $\rmC_\vtheta$ are $\rlambda_1 = \lambda_1 - n, \lambda_2, \ldots, \lambda_n$, which follows because $\vone$ is the first eigenvector of both $\mC_\vtheta$ and $\rmC_\vtheta$. Then, 
\begin{equation*}
1 - \frac{n}{\lambda_1}  = \frac{\lambda_1 - n}{\lambda_1} = \frac{\rlambda_1}{\rlambda_1 +n},
\end{equation*}
where now the right hand side is free of cancellation error.

>>>>

The process outlined above typically requires a computational cost of $O(N_{\text{opt}}n^3)$, where $N_{\text{opt}}$ is the number of optimization steps required to identify the hyperparameters. Our innovation is to pair low discrepancy nodes with matching covariance kernels that lower the  computational cost to $O(N_{\text{opt}} n \log n)$.   This approach is demonstrated using rank-1 lattice sequences and shift-invariant kernels.  Our algorithm is  implemented in the Guaranteed Automatic Integration Library (GAIL).

@@

Section \ref{sec:BC} explains the Bayesian approach to estimating the posterior cubature error and defines our automatic Bayesian cubature. Although much of this material is known, it is included for completeness.  We end Section \ref{sec:BC}  by demonstrating why Bayesian cubature is typically computationally expensive.
Section \ref{sec:fast_BC}  introduces the concept of covariance kernels that match the nodes, which expedites the computations required by our automatic Bayesian cubature. 
Section \ref{sec:shift_invariant_kernel} implements this concept for shift-invariant kernels and rank-1 lattice nodes.  This section also describes how to avoid cancellation error for covariance kernels of product form.  Numerical examples are provided in Section \ref{sec:NumExp} to demonstrate our new algorithm.  We conclude with a brief discussion.

@@

We use the following parameter values in the simulation: 
\begin{align*}
d' = 3, \quad \va = \begin{pmatrix}
-6 \\ -2 \\ -2
\end{pmatrix}, \quad 
\vb = \begin{pmatrix}
5 \\ 2 \\ 1
\end{pmatrix} , \quad 
\mL = \begin{pmatrix}
4 & 1 & 1 \\ 0 & 1 & 0.5 \\ 0 & 0 & 0.25
\end{pmatrix}.
\end{align*}
The node sets are randomly scrambled Sobol' points \cite{DicEtal14a,DicPil10a}.  The results are for 400 randomly chosen $\varepsilon$ in the interval $[10^{-5}, 10^{-2}]$ as shown in \figref{fig:MVN_Metern_d2b2}. In each run, the nodes are randomly scrambled.  The empirical Bayes credible intervals are used for stopping criteria.  We  observe that the algorithm meets the error criterion 95\% of the time even though we used 99\% credible intervals. One possible explanation is that the matrix inversions in the algorithm are ill-conditioned leading to numerical inaccuracies.  Another possible explanation is that this Mat\'ern covariance kernel is not a good match for the integrand.
%with Intel i7 3630QM and 16GB RAM memory
%On our test computer,

As shown in \figref{fig:MVN_Metern_d2b2}, the computation time increases rapidly with $n$. 
The  empirical Bayes estimation of $\vtheta$, which requires repeated evaluation of the objective function, is the most time consuming of all. It takes tens of seconds to compute $\hmu_n$ with $\varepsilon = 10^{-5}$.   In contrast, this example in Section \ref{sec:NumExp} take less than a hundredth of a second to compute $\hmu_n$ with the same $\varepsilon$ using our new algorithm. Not only is the Bayesian cubature with the Mat\'ern kernel slow, but also $\mC_\vtheta$ becomes highly ill-conditioned as $n$ increases.
So, Algorithm \ref{algorithm1}  in its current form is impractical when $n$ must be large. 
%\FJHNote{Jags, please check this re-write} %\JRNote{looks good}

\begin{figure}
	\centering
		\includegraphics[width=0.95\linewidth]{MVN_guaranteed_time_Matern_d2_2019-Jun-29}
	\centering
		\includegraphics[width=0.95\linewidth]{MVN_rapid_n_vs_time_Matern_d2_2019-Jun-29}
	\caption{Multivariate Gaussian probability estimated using Mat\'ern kernel with $d=2$ using the empirical Bayes stopping criterion. Top: Ratio of the integration error to the error tolerance versus execution time. Bottom: Execution time rapidly increases with increasing $n$. These figures can be reproduced using \code{matern\_guaranteed\_plots.m} in GAIL.}
	\label{fig:MVN_Metern_d2b2}
\end{figure}

@@

The generic automatic Bayesian cubature algorithm described in the previous section requires $\Order\bigl(n \$(f) + \linebreak[4] N_{\opt} [n^2 \$(C_\vtheta)  + n^3]\bigr)$ operations to compute the cubature. Now we explain how to speed up the calculations. A key is to choose covariance kernels that match the nodes, $\{\vx_i\}_{i=1}^n$, so that the vector-matrix operations required by Bayesian cubature can be accomplished using fast transforms at a computational cost of $\Order\bigl(n \$(f) + N_{\opt} [n \$(C_\vtheta)  + n \log(n)] \bigr)$.

@@

We make three assumptions that facilitate  fast computation:
\begin{subequations} \label{fastcompAssump}
	\begin{gather}
	\label{fastcompAssumpA}
	\mV \text{ may be identified analytically}, \\
	\label{fastcompAssumpB}
	\vv_1 = \vV_1 = \vone, \\
	\label{fastcompAssumpC}
	\mV^H \vb  \text{ requires only $\Order(n \log(n))$ operations } \forall \vb.
	\end{gather}
\end{subequations}
We call the transformation $\vb \mapsto \mV^H \vb$ a \emph{fast transform} and $C$ a \emph{fast transform kernel}.

@@

\begin{theorem} \label{thm:fastparam}
Under assumptions \eqref{fastcompAssump}, the parameters and credible interval widths in Theorem \ref{thm:param} may be expressed in terms of the fast transforms of the function data, the first column of the Gram matrix, and $\vc_\vtheta$ as follows:
\begin{align}
\nonumber
m_\MLE &=  m_{\full} = m_{\GCV} =  \frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\\
\nonumber
s^2_\MLE 
& =
\frac{1}{n^2} 
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}, \\
\nonumber
\widehat{\sigma}^2_{\textup{full}}
& =
\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}
\\
\nonumber
& \qquad \times
\left[\frac{\lambda_1}{n}{\left(1 - \frac{\widetilde{c}_1}{\lambda_1}\right)^2} + \left(c_0  - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i}\right) \right], \\
\nonumber 
s^2_{\textup{GCV}} & =  \frac 1{n} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1},
\end{align}
\begin{subequations}
	\label{fastTheta}
	\begin{align}
\nonumber
\vtheta_\MLE
&= 
\argmin_{\vtheta}
\left[
\log\left(
\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta, i}}
\right) \right .\\ 
\label{eqn_MLE_loss_func_optimized_2} %
& \hspace{3cm} \left .
 + \frac{1}{n}\sum_{i=1}^n \log(\lambda_{\vtheta,i})
\right],\\
\nonumber 
\vtheta_{\GCV} 
&= \argmin_\vtheta \left[ \log \left ( \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_{\vtheta, i}^2} 
\right) \right .\\
\label{thetaGCV}
& \qquad \qquad \qquad \qquad  \left . -2\log\left( \sum_{i=1}^n \frac{1}{\lambda_{\vtheta, i}} \right)
\right], 
\end{align}
\end{subequations}
\begin{align}
\nonumber
\hmu_\MLE  &= \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} +
\frac 1n \sum_{i=2}^n \frac{ \widetilde{c}_i^* \widetilde{y}_i}{\lambda_i}, \\
\nonumber
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
}, \\
\nonumber
\err_{\full} & = t_{n-1,0.995} \hsigma_{\textup{full}}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} \right.
\\ 
\nonumber
&\qquad \qquad \left . \times
\left( c_0 - \frac 1n \sum_{i=1}^n \frac{\abs{\widetilde{c}_i}^2}{\lambda_i} \right) 
\right\}^{1/2}.
\end{align}
Under the further assumption \eqref{addAssump}, it follows that 
\begin{equation}
\label{muhatGCV-FB-MLE-Simple}
\hmu_\MLE  = \hmu_{\full} = \hmu_{\GCV} =
\frac{\widetilde{y}_1}{n} = \frac 1n \sum_{i=1}^n y_i,
\end{equation}
and so $\hmu$ is simply the sample mean.  Also, under assumption \eqref{addAssump}, the credible interval widths simplify to
\begin{align}
\label{eq:errMLEAllAsump}
\err_\MLE  &
=
\frac{2.58}{n}\sqrt{
	\sum_{i=2}^{n} \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  
	\,
	\left( 1 -  \frac{n}{\lambda_1} \right) 
}, \\
\label{FJH:eq:errFullSimple}
\err_{\textup{full}}
&=
t_{n-1,0.995}
\sqrt{\frac{1}{n(n-1)} \sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i}  \left(\frac{\lambda_1}{n}  - 1  \right)}, \\
\nonumber
\err_{\textup{GCV}} & =
\frac{2.58}{n}\left\{\sum_{i=2}^n \frac{\abs{\widetilde{y}_i}^2}{\lambda_i^2}  \left [ \frac 1n \sum_{i=1}^n \frac{1}{\lambda_i} \right]^{-1} 
\right. \\ &\qquad \qquad \left . \times
\left( 1 -  \frac{n}{\lambda_1} \right)  
\right\}^{1/2}. \label{errGCVSimple}
\end{align}
In the formulas for the credible interval widths, $\vlambda$ depends on $\vtheta$, and  $\vtheta$ is assumed to take on the values $\vtheta_{\MLE}$ or $\vtheta_{\GCV}$ as appropriate.
\end{theorem}

@@

The preceding sections lay out an automatic Bayesian cubature algorithm whose computational cost is drastically reduced.  However, this algorithm relies on covariance kernel functions, $C_\vtheta$ and node sets, $\{\vx_i\}_{i=1}^n$ that satisfy assumptions \eqref{fastcompAssump}.  We  also want to satisfy assumption \eqref{addAssump}.  
To conveniently facilitate the fast transform, it is assumed in this section and the next that $n$ is power of $2$.  
%To facilitate the fast transform, $n$ must be power of $2$.

@@

The set of nodes used is defined by a shifted extensible integration lattice node sequence, which takes the form
\begin{equation} \label{eqn:lattice_def}
\vx_{i} = \vh \phi(i-1) + \vDelta \mod \vone, \qquad i \in \naturals.
\end{equation} 
Here, $\vh$ is a $d$-dimensional generating vector of positive integers, $\vDelta$ is some point in $[0,1)^d$, often chosen at random, and $\{\phi(i)\}_{i=0}^\infty$ is the van der Corput sequence, defined by reflecting the binary digits of the integer about the decimal point, i.e., 
\begin{align} \label{vdCDef}
\begin{array}{r|ccccccccccccc}
i & 0 & 1 & 2 & 3 & 4 &  5 & 6 & 7 & \cdots \\
i & 0_2 & 1_2 & 10_2 & 11_2 & 100_2 & 101_2 & 110_2 & 111_2  & \cdots\\
\toprule
\phi(i) & {}_2.0 &  {}_2.1 & {}_2.01 &  {}_2.11  & {}_2.001 &  {}_2.101 & {}_2.011 &  {}_2.111 & \cdots\\
\phi(i) & 0 &  0.5 &  0.25 & 0.75 &  0.125 & 0.625  &  0.375 & 0.875 & \cdots
\end{array}
\end{align}
Note that 
\begin{multline} \label{phiprop}
n\phi:\{0, \ldots, n-1 \} \to \{0, \ldots, n-1\} \\
\text{is one-to-one},
\end{multline}
assuming $n$ is a power of $2$.

An example of $64$ nodes is given in \figref{latticefig}.  The even coverage of the unit cube is ensured by a well chosen generating vector, $\vh$.  The choice of generating vector is typically done offline by computer search.  See \cite{DicEtal14a,HicNie03a} for more on extensible integration lattices.
\begin{figure}[htp]
	\centering
	\includegraphics[height=5cm]{ShiftedLatticePoints}
	\caption{Example of a shifted integration lattice node set  in $d=2$. 
	This figure can be reproduced using \code{PlotPoints.m} in GAIL} \label{latticefig}
\end{figure}

\subsection{Shift Invariant Kernels}
The covariance functions $C_\vtheta$ that match integration lattice node sets have the form
\begin{subequations} \label{eqn:ourkernel}
\begin{align} \label{eq:shInv}
C_\vtheta(\vt,\vx) = K_\vtheta(\vt - \vx \bmod \vone).
\end{align}
This is called a \emph{shift-invariant kernel} because shifting both arguments of the covariance function by the same amount leaves the value unchanged.  Here, $K_\vtheta$ is periodic and must  be of the form that ensures that $C_\vtheta$ is symmetric and positive definite, as assumed in \eqref{FJH:eq:CondPosDef}.

A family of shift-invariant kernels is constructed via even degree Bernoulli polynomials:
\begin{multline}
\label{the_kernel_eqn_bernoulli}
K_\vtheta(\vx) =
\prod_{l=1}^d \biggl[
1 - (-1)^{r} \eta B_{2r}( x_l) \biggr], \\  
\forall \vt,\vx \in [0,1]^d, \  \vtheta = (r,\eta), \ r \in \naturals, \ \eta > 0.
\end{multline}
\end{subequations}
Symmetric, periodic, positive definite kernels of this form appear in  \cite{DicEtal14a,Hic96a}.  Bernoulli polynomials are described in \cite[Chapter 24]{OlvEtal10a}.

Larger $r$ implies a greater degree of smoothness of the covariance kernel.  Larger $\eta$ implies greater fluctuations of the output with respect to the input.  Plots of $C_\vtheta(\cdot, 0.3)$ are given in \figref{fig:fourierkernel-dim1} for various $\vtheta=(r, \eta)$ values.

Lattice cubature rules are known to have convergence rates that depend on the smoothness of the integrands, but that are rather independent of the choice of the integration lattice \cite{DicEtal14a}.  Thus, we expect integration lattice node sets to perform well regardless of the smoothness of the covariance kernel.  The bigger concern is whether the derivatives of the integrand are as smooth as the covariance kernel implies.  This topic is touched upon again in Section  \ref{period_var_tx}.

@@

\subsection{The Gram Matrix as the Permutation of a Circulant Matrix}
For general shift-invariant covariance functions, the Gram matrix takes the form of a permutation of the rows and columns of a circulant matrix. By the properties of $\phi$ in \eqref{phiprop}, it follows that
\begin{equation} \label{PermMat}
    \mP = \bigl( \delta_{n\phi(i-1), j-1}  \bigr)_{i,j=1}^n
\end{equation}
is a permutation matrix, where $\delta_{\cdot,\cdot}$ is the Kronecker delta function.  Then,
\begin{align}
\nonumber
\mC_\vtheta &= \bigl ( C_\vtheta(\vx_i, \vx_j) \bigr)_{i, j = 1}^n \\
\nonumber
& = \Bigl ( K_\vtheta \bigl(\vh(\phi(i-1) - \phi(j-1) \bigr) \bmod \vone ) \Bigr)_{i, j = 1}^n \\
\nonumber
& \qquad \qquad  \text{by \eqref{eqn:lattice_def} and \eqref{eq:shInv}}  \\
\nonumber
& = \biggl( 
\sum_{i',j'=1}^n \delta_{n\phi(i-1), i'-1}  \,
K_\vtheta \bigl (\vh (i'-j')/n \bmod \vone \bigr )\\
\nonumber
& \qquad \qquad \times \delta_{j'-1,n\phi(j-1)} 
\biggr)_{i,j=1}^n \\
& = \mP \mK_{\vtheta}  \mP^T,  \label{shInvKernGramMatrix} 
\intertext{where } 
\mK_{\vtheta} &= \bigl ( K_\vtheta \bigl (\vh (i-j)/n \bmod \vone \bigr ) \bigr)_{i, j = 1}^n.
\end{align}

@@

Assumptions \eqref{fastcompAssumpA} and \eqref{fastcompAssumpB} can be verified by \eqref{PermMat}, \eqref{Keig}, and \eqref{Clateig}.  Assumption \eqref{fastcompAssumpC} is satisfied because $\mV^H \vb = \mW^H \mP^T \vb$ is just the discrete Fourier transform of a vector whose rows have been permuted.  This can be performed in $\Order(n \log(n))$ operations by the FFT algorithm. A proper scaling of the kernel $K_\vtheta$ ensures that assumption \eqref{addAssump} is satisfied.

@@

Any shift-invariant covariance kernel can be written as $C_\vtheta = 1 + \rC_\vtheta$, where $\rC_\vtheta$ is also symmetric and positive definite. The associated Gram matrix for $\rC_\vtheta$ is then $\rmC_\vtheta = \mC_\vtheta - \vone \vone^T$, and the eigenvalues of $\rmC_\vtheta$ are $\rlambda_1 = \lambda_1 - n, \lambda_2, \ldots, \lambda_n$, which follows because $\vone$ is the first eigenvector of both $\mC_\vtheta$ and $\rmC_\vtheta$. Then, 
\begin{equation*}
1 - \frac{n}{\lambda_1}  = \frac{\lambda_1 - n}{\lambda_1} = \frac{\rlambda_1}{\rlambda_1 +n},
\end{equation*}
where now the right hand side is free of cancellation error.
